---
title: "Using Provided Data for Business Objectives"
author: "Rachael Sanderson"
date: "19 November 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
library(ProjectTemplate)

```

```{r load_project, include = FALSE}
load.project()

```
##The Focus of Analysis

- The Goals of Learning Analytics

- Selecting the Data
    - Quiz Questions
    - Videos

- Respresenting Findings
    - Does the presentation of data influence its success in meeting the business goals?
    
##Analysing Attainment: The Percentage of Correct Answers by Cohort
Option 1: Scatter Charts - Seeing each correct answer

```{R CorrectGraph, fig.width = 8, echo = FALSE}
Correct = ggplot(data=CorrectDF, x=Question, Y = Correct_Answers)
C1 = Correct + geom_point(aes(x=Question, y=Course1, colour = "Course1"))
C2 = C1 + geom_point(aes(x=Question, y=Course2, colour = "Course2"))
C3 = C2 + geom_point(aes(x=Question, y=Course3, colour = "Course3"))
C4 = C3 + geom_point(aes(x=Question, y=Course4, colour = "Course4"))
C5 = C4 + geom_point(aes(x=Question, y=Course5, colour = "Course5"))
C6 = C5 + geom_point(aes(x=Question, y=Course6, colour = "Course6"))
C7 = C6 + geom_point(aes(x=Question, y=Course7, colour = "Course7"))
C8 = C7 +labs(title="The Number of Correct Answers for Each Question",
        x ="Question", y = "Number of Correct Answers") + theme(axis.text.x=element_blank())

C8
```

##Analysing Attainment: The Percentage of Correct Answers by Cohort
Option 2: BoxPlots - Summarising the course

```{R Question_Box, echo= FALSE, fig.width=8}
Correct = within(CorrectDF, rm(Question, Average))
CorrectVec = as.vector(as.matrix(Correct))
Course = rep(1:7, each = 22)
BoxDF = data.frame(Course = Course, Score = CorrectVec)

Box = ggplot(data = BoxDF, x = Course, y = Score)
B1 = Box + geom_boxplot(aes(group = Course, x = Course, y = Score, fill = factor(Course)))
B2 = B1 + labs(title="The Average Number of Correct Answers Throughout a Course Run",
        x ="Course", y = "Percentage of Correct Answers (%)")

B2
```

 - Overall reflections
    - Findings
    - Visuals

##The Use of Videos: How many views did each video have?

``` {R Views_Dataframe, echo = FALSE}
DFViews

```

##The Use of Videos: Scatter & Line

```{R Scatter, include = FALSE}
Graph=ggplot (data = DFViews, aes (x = Step, y= Total.Video.Views))
g3 = Graph + geom_point(aes(x = Step, y=Video3, colour = "Course Run 3")) #adding each course run
g4 = g3+ geom_point (aes (x=Step, y=Video4, colour = "Course Run 4"))
g5 = g4+ geom_point (aes (x=Step, y=Video5, colour = "Course Run 5")) 
g6 = g5+ geom_point (aes (x=Step, y=Video6, colour = "Course Run 6"))
g7= g6+ geom_point (aes (x=Step, y=Video7, colour = "Course Run 7"))

g8 = g7 + geom_line (aes(x=Step, y=Mean, colour = "Average Views")) #adding the average line
g9 = g8 +labs(title="Total Viewings of each Video",
        x ="Video Step", y = "Total Views")
g9
```

```{R Line, include = FALSE}
Line=ggplot (data = DFViews, aes (x = Step, y= Views))
L3 = Line + geom_line(aes(x = Step, y=Video3, colour = "Course Run 3")) #adding each run
L4 = L3+ geom_line (aes (x=Step, y=Video4, colour = "Course Run 4"))
L5 = L4+ geom_line (aes (x=Step, y=Video5, colour = "Course Run 5")) 
L6 = L5+ geom_line (aes (x=Step, y=Video6, colour = "Course Run 6"))
L7= L6+ geom_line (aes (x=Step, y=Video7, colour = "Course Run 7"))
L8 = L7 + geom_line (aes(x=Step, y=Mean, colour = "Average Views")) #adding the average line
L9 = L8 +labs(title="Total Viewings of each Video",
        x ="Video Step", y = "Total Views")
L9

```

```{R Both, echo = FALSE, fig.width = 8}
grid.arrange(g9,L9, ncol = 2)

```

##The Use of Videos: Proportions
```{R Proportion, echo = FALSE}

Line=ggplot (data = DFViewsProportion, aes (x = Step, y= Video3)) #initial line coordinates

L3 = Line + geom_point(aes(x = Step, y=Video3, colour = "Course Run 3")) #adding each run
L4 = L3+ geom_point (aes (x=Step, y=Video4, colour = "Course Run 4"))
L5 = L4+ geom_point (aes (x=Step, y=Video5, colour = "Course Run 5")) 
L6 = L5+ geom_point (aes (x=Step, y=Video6, colour = "Course Run 6"))
L7= L6 + geom_point (aes (x=Step, y=Video7, colour = "Course Run 7"))
L8 = L7 + geom_line (aes ( x=Step, y=DFViewsProportion$Average, colour = "Average"))

L8


```

##Merging the Videos and Questions

- Comparing Steps
```{R Question_Lines, echo = FALSE, warning = FALSE}
 StepActivity = cyber.security.7.question.response
Questions = unique(StepActivity$quiz_question, incomparables = FALSE)
  
Steps=c((StepActivity$week_number)+((StepActivity$step_number)/100)) #acknowledging data complication
  StepActivity = cbind(StepActivity, Steps) #editing the data with the new column
  S= unique(StepActivity$Steps, incomparables = FALSE) #list the unique steps
  
Q9 = L8 + geom_vline(xintercept = (S[1]), (aes (colour = "red")))
Q10 = Q9 + geom_vline(xintercept = (S[2]), (aes (colour = "red")))
Q11 = Q10 + geom_vline(xintercept = (S[3]), (aes (color = "red")))
Q12 = Q11 + geom_vline(xintercept = (S[4]), (aes (color = "red")))
Q13 = Q12 + geom_vline(xintercept = (S[5]), (aes (color = "red")))

Q13

```

##Conclusion
- Reflections: The observed trends of attainment and video usage 

- Are the findings effectively translated?
